Ontology Matching with word2vec

### 2016-05-14 Paper/Draft.tex
Corrections up to p.9

 \pagenumbering{arabic}
 \section{Introduction}
 It is a Friday afternoon and you realise you want to watch a movie with some friends this weekend. So, what do you do? To achieve this goal you must take four steps.

@@

 On the Semantic Web, Ontologies are used to define and share knowledge. The term ontology will be properly defined later on, but for now it suffices to know that an ontology is a set of rules that determine what relations are allowed and required in a given domain of discourse. Without ontologies, agents on the Semantic Web would be unable to communicate, as all concepts are defined in an Ontology.

 However, using one single ontology is infeasible. This is due to the fact that inference and search time scales with Ontology size, and the number of concepts that are needed on the Semantic Web is enormous. Every single concept that is present in the billions of Terabytes of text, images and sounds that are on the web needs to be represented, for every professional domain and science, in every language. And all these concepts are related. Searching through this Ontology just to find one concept would be incredibly costly. And since every interaction on the Semantic Web requires lookups in an Ontology or corresponding Knowledge Base, these lookups need to be fast.

 The solution to this problem is to split this Ontology into many smaller Ontologies. However, as often is the case in Computer Science, this creates two new, albeit smaller, problems.

 The first problem is the problem of splitting the Ontology. How do we find the distribution of all concepts over the different Ontologies to maximize their usefulness. Usefulness in this case should take into account co-occurrence of concepts. Naturally, connected concepts should be more likely to occur in the same Ontology as concepts that are not connected, since it is likely that agents that want to know about a concept also want to know things about its neighbours.

 The second problem, the one I address in this thesis, is the problem of conversations that cover multiple domains. These conversations will occur more often if there are more Ontologies, but they can always occur if the number of Ontologies is larger than one. When a conversation requires information from multiple domains, the Ontologies corresponding to those domains need to be merged. This means connecting the related concepts from the different Ontologies. Some research has focused on ontology merging, with promising but slowing results.

 The problem of Ontologies on the Semantic Web is to find those Ontologies that are as small as possible to improve search speed, but need to be merged as little as possible, to ensure as little time is used to merge the Ontologies before they can be searched. If it is possible to quickly and reliably merge Ontologies, Ontologies can be made small and communication on the Semantic Web can be done quickly.

 Recently, a model that embeds words in a vector space that appears to contains semantics has become popular. Since it is unsupervised and can be used to calculate a semantic distance between words, it may very well be applicable to ontology matching.

@@

\newpage
\section{Context}
 In this section I will discuss all concepts that are required to understand the study. Firstly, the Semantic Web, the key technology that depends very strongly on Ontology matching, will be discussed. Then the concept of the Ontology itself will be addressed. Lastly, current alignment strategies, their problems and other research and technologies will be discussed.
 \subsection{The Semantic Web}
 The Semantic Web, also called Web 3.0, is a concept of a Web that enhances the User Experience in ways that the current Web can not provide. Web 2.0, the current Web, models the internet as \textit{hubs} (or sites) with a specific purpose and limited interconnectivity. For example, Facebook is a hub for sharing content with people you consider friends, but if you want to share that content with more people you will have to go to another hub, like imgur or Pinterest. It is possible to link to content from other hubs using URLs, but this moves the user to another hub, rather than connecting the content present in the hubs.
The Semantic Web, on the other hand, contains sources of \textit{information} that can be combined whenever the user needs it. To stick with the social media example, a person can create a piece of content, say a picture of a tree, en shares it with his friends. One of his friends then can share this piece of content with a group of people that likes nature pictures. However, everyone can access all references someone makes. So if a person from the nature group notices the tree is ill and comments on this, the original poster can see this comment and act accordingly. On the current web the friend who shared the picture with the group has to be contacted directly by the commenter and then has to relay the message manually.

@@

 All of this is possible if the softbots can access all of the information reliably. In the current Web, this is not possible, as all information is written in an ambiguous, unstructured format called human language. To allow softbots to access the same information that we can, the Semantic Web proposes to add a layer to the internet where all information is saved in a unified format that is unambiguous, robot-interpretable and can be used to store any type of knowledge.

@@

 To summarize, the goal of the Semantic Web is to connect all online information to help humans. Only when all information can be connected will it be possible to aid humans in their needs. But humans will only accept the Semantic Web if it is useful to them, so the Semantic Web must show its applications and usefulness before it will be used in daily life.
 \subsubsection{Current state}
 Currently, people are not aware of the Semantic Web. Either websites and applications do not use any semantic information, or it is used on the back-end, hidden with other complex programs that users will not understand. So what is the Semantic Web used for nowadays, if at all? In this section, I will list a number of general ways people use Semantic Web technologies as well as specific applications. Note that these are Semantic Web technologies, i.e. technologies that came into existence through Semantic Web research and can eventually be used in the Semantic Web, but currently are not used for the Semantic Web since it does not exist yet.

 The Semantic Web is currently divided into two groups: the people who need high accuracy and the people who need high recall. The first group consist of scientists that work on specific fields that require exact knowledge, whereas the second group are data scientists that want big amounts of data available. In the past it was not possible to extract information from texts and other sources automatically, so

 The group that requires high recall use hand-crafted knowledge bases, created by experts, containing only facts that are accurate, verified and relevant. This method of creating a knowledge base does not scale very well and usually consists of fewer than ten thousand facts. Because of this, the knowledge bases are restricted to small domains, including some parts of medicine and cultural heritage. \cite{byrne, le2002}

 The group that requires a high recall, i.e. requires a huge amount of data, automated methods are required. These methods work fast, parsing thousands of websites a minute, but are inclined to make mistakes. Often, these methods use crowd-sourcing to improve accuracy, using experts or the general public to suggest or check facts. \cite{zhou2007}

@@

 Large ontologies already exist, with the largest, DBPedia, boasting almost two billion triples and the three hundred combined Large Open Data databases contain over thirty billion. \cite{dbpedia, lod} The LOD is simply one huge alignment of hundreds of knowledge bases... %Doorschrijven

@@

 \subsection{Ontologies}
 Earlier, ontologies were referred to as 'a set of rules that determine what relations are allowed and required.' Now, the term will be defined more extensively, the standard format will be described and the uses of ontologies will be discussed.

 An ontology describes concepts as their relationships to other concepts. The concepts that are described determine the domain the ontology covers, and the relationships determine the rules the ontology imposes on the domain. For example, a hierarchy is an ontology that describes concepts that are subclasses of other concepts. Therefore the relationships are 'part-of' relationships: birds are part-of the animals concept class.
 Concepts can be instantiated, which is done in a Knowledge Base (KB). KBs and Ontologies are often confused, and are indeed very similar in structure. The Knowledge Base in our example can contain actually existing birds or fictional birds, and refers to the Ontology concept of bird to embed the instantiations with meaning.

 \subsubsection{OWL}
 OWL is a family of languages and syntaxes that can be used to create an ontology. The different languages are designed around requirements of possible relationships and concept definitions, and thus may differ a lot between them. The W3C has defined three variants which are the bases for all other adaptations. The variants trade off levels of expressiveness versus computability.
 \subsubsection{SPARQL}
 SPARQL is the query language for knowledge bases and ontologies. It allows a user to select nodes whose context match the query, similar to how rows are selected in SQL based on the row contents.
 \subsubsection{Role}
 %Beschrijf de drie manieren om top-down ontologien te maken. Geen intra-ontologiale verwijzingen, verwijzingen of overlap (maar dan nog steeds verwijzingen). Beschrijf bottom-up ontologie maken.

 The reason ontologies are used in databases consists of multiple parts. Firstly, it allows querying languages to optimize the search process. Secondly, it allows expansion of information on a node. Thirdly, it allows for consistency checking a Knowledge Base. Lastly, it allows merging of knowledge bases.

@@

 When two knowledge bases need to be merged, aligned or matched, instantiations that refer to the same thing need to be merged. For example, if two person databases both contain references to the same person, all information on that person needs to be linked to the same object, such that information of that person from both knowledge bases can be combined. Merging the ontologies of the two knowledge bases vastly improves the alignment between the two knowledge bases. However, ontology and knowledge base merging is not a trivial problem and actually the problem that this thesis addresses. Therefore it will be explained more in-depth later on.
 \subsubsection{Aristotle and Plato}
 All ontologies lie on a spectrum of formality. On the one end of the spectrum are the Platonic ontologies, which should only contain concepts that are perfectly defined and constraints which are completely binding. The name refers to the Platonic philosophy that all concepts are existing entities and can thus be perfectly captured in a definition.

 On the other end of the spectrum are the Aristotelian ontologies, which do not necessarily contain perfectly defined concepts, but rather concepts as we observe them. Aristotle disagreed with Plato's theory that concepts are existing things, and thus these ontologies are named after him.

@@

 The advantage of a Platonic ontology would be the fact that every query would result in a fact, since the ontology itself is perfect. The disadvantage is that creating such an ontology is much more costly than an Aristotelian ontology, since every concept and constraint needs to be correct in all cases.

@@

 Platonic ontologies are often used in areas where precision is important, like medicine, where lives depend on the information contained in the ontology and its knowledge base. Aristotelian ontologies are more common on the Semantic Web, where massive amounts of data need to be represented and queried.
 \subsection{Current alignment strategies} \label{strategies}
 Many different alignment strategies have already been developed. All strategies follow the same two-step approach. The two steps are independent and as such different methods can be used interchangeably. The first step is to generate an initial correspondence set, where correspondences between nodes are found based on just their labels and meta-data. The second step is to use this initial set and the structure of the ontology to find more correspondences. These steps are called the terminological and the structural steps. Some strategies also use extra steps like the extensional and semantic steps, which use vector space models and inference, respectively. https://hal.inria.fr/hal-00917910/document

 Popular terminological strategies are WordNet comparison and edit distance. The former uses the popular WordNet hierarchy as a distance measure between two concepts, e.g. the number of edges between the concepts and their least general common superconcept. \cite{lin2008}
 Edit distance is a purely string-based similarity measure. The similarity is a (weighted) count of edits required to transform one word into another, where common edits are insertion, deletion and replacement. Similar methods include substring matching and n-gram matching, which compare parts of the strings to find similarities.\cite{singh2014, levenshtein}
 \subsection{Problems in ontology alignment}
 There are a number of different problems ... % Opsommen en samenvatten

@@

 Recently, an effective and efficient skip-gram model has been developed, called word2vec, which uses a number of extensions over previous methods that enable it to efficiently learn an effective representation. /cite{w2v}
 Word2vec learns this representation by trying to predict the context of a word. First the vocabulary is one-hot encoded. Then the model learns to predict a context vector that minimizes the distance to the context one-hot encoded vectors. The model is a neural network with one hidden layer. Since the input is one-hot encoded, every word in the vocabulary is represented by one row in the weight matrix between the input layer and hidden layer. This row is the vector representation of the word.

 Word2vec has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on. \cite{ling2015}
 Another interesting development is the multi-sense word2vec, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment.\cite{multisense}
 \subsection{Used technology}
 \subsubsection{AgreementMaker} \label{agreementmaker}
 AgreementMaker is an ontology matching system that obtained the highest F-measure in 6 of the 7 ontology matching tracks of OAEI 2015. \cite{OAEI2015} The system has also been updated the most recent at the time of writing\cite{agreementmaker}.

>>>>

 \pagenumbering{arabic}
 \section{Introduction} \label{intro}
 It is a Friday afternoon and you realise you want to watch a movie with some friends this weekend. So, what do you do? To achieve this goal you must take four steps.

@@

 On the Semantic Web, Ontologies are used to define and share knowledge. The term ontology will be properly defined later on, but for now it suffices to know that an ontology is a set of rules that determine what relations are allowed and required in a given domain of discourse. As a consequence, if two agents want to communicate in a given domain, they need to agree on the rules of that domain, or in other words, share the ontology of that domain.

 However, it is unlikely that a conversation only covers one domain. To solve this, there are two potential solutions. Either you ensure that all possible combinations of domains are covered by an ontology, which means one ontology that covers all domains is required, or, alternatively, ontologies are combined on-the-fly.

 The first option, using one single ontology, is infeasible. This is due to the fact that inference and search time scales with Ontology size, and the number of concepts that are needed on the Semantic Web is enormous. Every single concept that is present in the billions of Terabytes of text, images and sounds that are on the web needs to be represented, for every professional domain and science, in every language. And all these concepts are related. Searching through this Ontology just to find one concept would be incredibly costly. And since every interaction on the Semantic Web requires lookups in an Ontology, these lookups need to be fast.

 The second option, the one I address in this thesis, is the problem of combining ontologies from multiple domains. Combining in this sense means connecting the related concepts from the different Ontologies. Some research has focused on Ontology Merging, with promising results. However, recently research has been slowing to the point where there is little improvement.

 One big problem in Ontology merging is the fact that in different domains, words can have different meanings. Common examples are synonyms like author and writer, homonyms like bank (for money or by a river) and ambiguity. Ambiguity means that similar words mean slightly different things, or a word might have multiple related meanings. There are a few ways to deal with these forms of ambiguity already in use, however they are not sufficient.

 A recently developed model for word embedding may be used to deal with ambiguity. It maps words to a vector space that has semantic properties, and thus can be used to calculate the relatedness of two words. There are a number of extensions that can deal with homonyms as well as data structured other than the continuous bag of words representation that the initial model was developed for. These facts show that it may be possible to make an embedding model for Ontologies and use the word relatedness to combine Ontologies.

@@

\newpage
\section{Context}
 In this section I will discuss all concepts that are required to understand the study. Firstly, the Semantic Web, the key technology that depends very strongly on Ontology matching, will be discussed. Then the concept of the Ontology itself will be addressed. Lastly, current alignment strategies, their problems and other research and technologies will be discussed.
 \subsection{The Semantic Web}
 The Semantic Web, also called Web 3.0, is a concept of a Web that enhances the User Experience in ways that the current Web can not provide. Web 2.0, the current Web, models the internet as \textit{hubs} (or sites) with a specific purpose and limited interconnectivity.\footnote{For example, Facebook is a hub for sharing content with people you consider friends, but if you want to share that content with more people you will have to go to another hub, like imgur or Pinterest.} It is possible to link to content from other hubs using URLs, but this moves the user to another hub, rather than connecting the content present in the hubs.
The Semantic Web, on the other hand, contains sources of \textit{information} that can be combined whenever the user needs it. To stick with the social media example, a person can create a piece of content, say a picture of a tree, and share it with his friends. One of his friends then can share this piece of content with a group of people that likes nature pictures. However, everyone can access all references someone makes. So if a person from the nature group notices the tree is ill and comments on this, the original poster can see this comment and act accordingly. On the current web the friend who shared the picture with the group has to be contacted directly by the commenter and then has to relay the message manually.

@@

 All of this is possible if the softbots can access all of the information reliably. In the current Web, this is not possible, as all information is written in an ambiguous, unstructured format called natural language. To allow softbots to access the same information that we can, the Semantic Web proposes to add a layer to the internet where all information is saved in a unified format that is unambiguous, robot-interpretable and can be used to store any type of knowledge.

@@

 To summarize, the goal of the Semantic Web is to connect all online information to help humans. Only when all information can be connected will it be possible to aid humans in their needs. But humans will only accept the Semantic Web if it is useful to them, so the Semantic Web must show its applications and usefulness before it will be used in daily life.
 \subsubsection{Current state} %TODO This bit can use some work
 %TODO Cite Peter Mika (see Arjen notes 1)
 Currently, people are not aware of the Semantic Web. Either websites and applications do not use any semantic information, or it is used on the back-end, hidden with other complex programs that users will not understand. So what is the Semantic Web used for nowadays, if at all? In this section, I will list a number of general ways people use Semantic Web technologies as well as specific applications. Note that these are Semantic Web technologies, i.e. technologies that came into existence through Semantic Web research and can eventually be used in the Semantic Web, but currently are not used for the Semantic Web since it does not exist yet.

 Semantic Web systems are currently divided into two groups: the systems that need high accuracy and the ones that need high recall. The first group are deployed in domains that require exact knowledge, whereas the second group are systems that must make available a big amount of data. In the past it was not possible to extract information from texts and other sources automatically, so all systems were hand-crafted and thus highly accurate but very small. This also impacted research, which did not focus on scaling ontologies until recently.

 The group that requires high recall use hand-crafted knowledge bases, created by experts, containing only facts that are accurate, verified and relevant. This method of creating a knowledge base does not scale very well and usually consists of fewer than ten thousand facts. Because of this, the knowledge bases are restricted to small domains, including some parts of medicine and cultural heritage\cite{byrne, le2002}.

 For the group that requires a high recall, or in other words a big amount of data, automated methods are required. These methods work fast, parsing thousands of websites a minute, but are inclined to make mistakes. Often, these methods use crowd-sourcing to improve accuracy, using experts or the general public to suggest or check facts\cite{zhou2007}.

@@

 Large ontologies already exist, with the largest, Freebase, boasting almost two billion triples and the three hundred combined Large Open Data databases contain over thirty billion\cite{dbpedia, lod}. The LOD is simply one huge alignment of hundreds of knowledge bases... %Doorschrijven

@@

 \subsection{Ontologies}
 Earlier, ontologies were referred to as `a set of rules that determine what relations are allowed and required.' Now, the term will be defined more extensively, the standard format will be described and the uses of ontologies will be discussed.

 An ontology describes concepts as their relationships to other concepts. The concepts that are described determine the domain the ontology covers, and the relationships determine the rules the ontology imposes on the domain. For example, a hierarchy is an ontology that describes concepts that are subclasses of other concepts. Therefore the relationships are 'is-a' relationships: a bird \emph{is an} animal.
 Concepts can be instantiated, which is done in a Knowledge Base (KB). KBs and Ontologies are often confused, and are indeed very similar in structure. The Knowledge Base in our example can contain actually existing birds or fictional birds, and refers to the Ontology concept of bird to embed the instantiations with meaning.

 \subsubsection{OWL}
 OWL is a family of languages and syntaxes that can be used to create an ontology. The different languages are designed around requirements of possible relationships and concept definitions, and thus may differ a lot between them. The W3C has defined three variants which are the bases for all other adaptations. The variants trade off levels of expressiveness versus computability. %TODO consider SKOS
 \subsubsection{SPARQL}
 SPARQL is the query language for knowledge bases and ontologies. It allows a user to select nodes whose context match the query, similar to how rows are selected in SQL based on the row contents.
 \subsubsection{Uses of Ontologies}
 %Beschrijf bottom-up ontologie maken.
 There are three main reasons Ontologies are used. Firstly, it allows querying languages to optimize the search process. Secondly, it allows expansion of information on a node. Thirdly, it allows for consistency checking a Knowledge Base. Lastly, it allows merging of knowledge bases.

@@

 When two knowledge bases need to be merged, aligned or matched, instantiations that refer to the same thing need to be merged. For example, if two person databases both contain references to the same person, all information on that person needs to be linked to the same object, such that information of that person from both knowledge bases can be combined. Merging the ontologies of the two knowledge bases vastly improves the alignment between the two knowledge bases. However, ontology and knowledge base merging is not a trivial problem and actually the problem that this thesis addresses. Therefore it will be explained more in-depth later on. %TODO add \ref
 \subsubsection{Aristotle and Plato}
 All ontologies lie on a spectrum of formality. On the one end of the spectrum are the Platonic ontologies, which should only contain concepts that are perfectly defined and constraints which are completely binding. The name refers to the Platonic philosophy that all concepts are existing entities and can thus be perfectly captured in a definition.

 On the other end of the spectrum are the Aristotelian ontologies, which do not necessarily contain perfectly defined concepts, but rather concepts as we observe them. Aristotle disagreed with Plato's theory that concepts are existing things, and thus these ontologies are named after him. %TODO cite

@@

 The advantage of a Platonic ontology would be the fact that every query would result in a fact, since the ontology itself is perfect. The disadvantage is that creating such an ontology is much more costly than an Aristotelian ontology, since every concept and constraint needs to be correct in all cases. %TODO cite Lakoff

@@

 Platonic ontologies are often used in areas where precision is important, like medicine, where lives depend on the information contained in the ontology and its knowledge base. Aristotelian ontologies are more common on the Semantic Web, where massive amounts of data need to be represented and queried.
 \subsection{Current alignment strategies} \label{strategies}
 Many different alignment strategies have already been developed. All strategies follow the same two-step approach. The two steps are independent and as such different methods can be used interchangeably. The first step is to generate an initial correspondence set, where correspondences between nodes are found based on just their labels and meta-data. The second step is to use this initial set and the structure of the ontology to find more correspondences. These steps are called the terminological and the structural steps. Some strategies also use extra steps like the extensional and semantic steps, which use vector space models and inference, respectively. %TODO cite https://hal.inria.fr/hal-00917910/document

 Popular terminological strategies are WordNet comparison and edit distance. The former uses the popular WordNet hierarchy as a distance measure between two concepts, e.g. the number of edges between the concepts and their least general common superconcept\cite{lin2008}.
 Edit distance is a purely string-based similarity measure. The similarity is a (weighted) count of edits required to transform one word into another, where common edits are insertion, deletion and replacement. Similar methods include substring matching and n-gram matching, which compare parts of the strings to find similarities\cite{singh2014, levenshtein}.
 \subsection{Problems in ontology alignment}
 There are a number of different problems ... % Opsommen en samenvatten

@@

 Recently, an effective and efficient skip-gram model has been developed, called word2vec, which uses a number of extensions over previous methods that enable it to efficiently learn an effective representation/cite{w2v}.
 Word2vec learns this representation by trying to predict the context of a word. First the vocabulary is one-hot encoded. Then the model learns to predict a context vector that minimizes the distance to the context one-hot encoded vectors. The model is a neural network with one hidden layer. Since the input is one-hot encoded, every word in the vocabulary is represented by one row in the weight matrix between the input layer and hidden layer. This row is the vector representation of the word.

 Word2vec has been expanded in a number of different ways. For example, word order can be preserved, leading to a similarity measure that is closer to syntax, as syntax defines which words go where in a sentence. This preservation was done by having multiple output layers, one for every word in the context. The research shows that the input does not have to be a bag-of-words representation, which will be useful later on\cite{ling2015}.
 Another interesting development is the multi-sense word2vec, which allows for multiple vector representations per word, depending on the number of different definitions a word has. For example, the word bank represents both the monetary institute and the riverside, which would both have a different representation in this extension. It can differentiate between two different meanings based on the context. This will help with resolving ambiguity, a very important factor in ontology alignment\cite{multisense}.
 \subsection{Used technology}
 \subsubsection{AgreementMaker} \label{agreementmaker}
 AgreementMaker is an ontology matching system that obtained the highest F-measure in 6 of the 7 ontology matching tracks of OAEI 2015 \cite{OAEI2015}. The system has also been updated the most recent at the time of writing\cite{agreementmaker}.
